Timeliness of security updates for Android

### 2014-10-30 securityupdates.tex
More tweaks from Alastair's comments

\label{sec:android_update_process}
\begin{figure}
 \centering
 \def\svgwidth{\columnwidth}
 \import{figures/}{update_ecosystem.pdf_tex}
 \caption{Flow of updates between participants in the Android ecosystem.
 Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.
 Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.}
 \label{fig:update_ecosystem}
\end{figure}
To understand how vulnerabilities in Android are fixed we must examine the Android update process which we model in Figure~\ref{fig:update_ecosystem}.
There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and so can include any compatible versions of those projects, including those which fix security vulnerabilities in them.
Android also incorporates various drivers for different bits of hardware.
The Android platform is then built on top of those with kernel and userspace components by Google.
The code for each update is kept secret\footnote{\url{https://source.android.com/source/code-lines.html}}\todolater{Can we quantify this keeping the code secret? Is it worth it?} until after the update has been published.
Manufacturers may receive advanced copies in order to prepare handsets.
Then the Manufacturers take the update and customise it before passing it on to the operator.
The Operator may then make further customisations and do further testing before shipping the update to the Device.
Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google are the same entity (at least for the purpose of a particular phone, such as with Nexus phones).
Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android.
\todolater{use statistics from samsung-updates.com -> how many binaries are there per device?}

The numbers of Devices (\daNumOSDevices), Operators (\daNumOperators) and Manufacturers (\daNumManufacturers) in Figure~\ref{fig:update_ecosystem} come from the \da\ data.
Manufacturer and Operator counts were obtained by normalising the results reported by Android to \da\ of the device manufacturer and active network operator.
This normalisation involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `lg'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some manufacturers and operators which are not included in the \da\ data.

@@

\subsubsection{Results}
\begin{figure}
 \centering
 \begin{subfigure}[b]{\columnwidth}
  \includegraphics[width=\columnwidth]{figures/device-data-all-security}
  \caption{Number of vulnerabilities affecting each device over time}
  \label{fig:device_data_security}
 \end{subfigure}
 \begin{subfigure}[b]{\columnwidth}
  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
  \caption{OS versions over time}
  \label{fig:device_data_os}
 \end{subfigure}
 \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data}
\end{figure}
The logitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data_security}.
The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
This implies that these devices had been abandoned and were not receiving updates which is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
Some devices start off in 2011 exposed to known security vulnerabilities and are still exposed to those and additional ones in 2014.

@@

By taking the number of updates observed and the number of device days of data we have collected we can compute the number of updates received by a device per year to be \daUpdatesPerYear.
This compares badly with the number of vulnerabilities discovered per year of \avoVulnsPerYear.

\todo{That comparison is not completely fair as not all of those vulnerabilites will affect all devices, similarly we may be counting device days for which we only have one day and hence could perhaps not observe an update.}

@@

\section*{Acknowledgements}
Thanks to David Robertson for helpful advice on statistical analysis.
Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case and Khilan Gudka for reporting vulnerabilities in Android.

>>>>

\label{sec:android_update_process}
\begin{figure}
 \centering
 \def\svgwidth{\columnwidth}
 \import{figures/}{update_ecosystem.pdf_tex}
 \caption{Flow of updates between participants in the Android ecosystem.
 Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.
 Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.}
 \label{fig:update_ecosystem}
\end{figure}
To understand how vulnerabilities in Android are fixed we must examine the Android update process which we model in Figure~\ref{fig:update_ecosystem}.
There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and so can include any compatible versions of those projects, including those which fix security vulnerabilities in them.
Android also incorporates various drivers for different bits of hardware.
The Android platform is then built on top of those with kernel and userspace components by Google.
The code for each update is kept secret\footnote{\url{https://source.android.com/source/code-lines.html}}\todolater{Can we quantify this keeping the code secret? Is it worth it?} until after the update has been published.
Manufacturers may receive advanced copies in order to prepare handsets.
Then the Manufacturers take the update and customise it before passing it on to the operator.
The Operator may then make further customisations and do further testing before shipping the update to the Device.
Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google collaborate closely for the purpose of a particular phone, such as with Nexus phones, and hence are harder to separate.
Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android~\cite{Wagner2013}.
\todolater{use statistics from samsung-updates.com -> how many binaries are there per device?}

The numbers of Devices (\daNumOSDevices), Operators (\daNumOperators) and Manufacturers (\daNumManufacturers) in Figure~\ref{fig:update_ecosystem} come from the \da\ data.
Manufacturer and Operator counts were obtained by normalising the results reported by Android to \da\ of the device manufacturer and active network operator.
This normalisation is a manual task involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `lg'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some manufacturers and operators which are not included in the \da\ data.

@@

\subsubsection{Results}
\begin{figure}
 \centering
 \begin{subfigure}[b]{\columnwidth}
  \includegraphics[width=\columnwidth]{figures/device-data-all-security}
  \caption{Number of vulnerabilities affecting each device over time}
  \label{fig:device_data_security}
 \end{subfigure}
 \begin{subfigure}[b]{\columnwidth}
  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
  \caption{OS versions over time}
  \label{fig:device_data_os}
 \end{subfigure}
 \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device.}
\end{figure}
The logitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data_security}.
The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
This implies that these devices had been abandoned and were not receiving updates which is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
Some devices start off in 2011 exposed to known security vulnerabilities and are still exposed to those and additional ones in 2014.

@@

By taking the number of updates observed and the number of device days of data we have collected we can compute the number of updates received by a device per year to be \daUpdatesPerYear.
This compares badly with the number of critical vulnerabilities discovered per year of between \avoVulnsPerYearAllAndroid\ (affecting all Android) and \avoVulnsPerYear\ (including the manufacturer specific ones in \avo).

@@

\section*{Acknowledgements}
Thanks to David Robertson for helpful advice on statistical analysis.
Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting vulnerabilities in Android.
